Leaf traits drive differences in biomass partitioning among major plant functional types

### 2015-07-01 manuscript.Rnw
minor edits to text

Overall, our results establish general patterns about plant construction and thus lay an empirical base against which models can be benchmarked. We found that PFTs differed in LMF, and that LMF was proportional to LMA. It thus seems possible to integrate a key leaf trait with whole-plant modelling of biomass partitioning. A recent study compared allocation routines in a number of leading ecosystem models \citep{dekauwe2014}, and recommended constraining allocation by observed biomass fractions instead of using constant allocation fractions. Based on our findings, a first approximation within GVMs would be to assume leaf area to stem cross-sectional area a parameter that does not vary between PFTs, and vary biomass allocation accordingly. LMA, already a parameter in most GVMs, then gives the ratio of leaf biomass to stem area. Some models already incorporate a similar algorithm (see Methods~\ref{sec:gvmsummary}), while other algorithms may be tuned to yield similar patterns between PFTs as we have presented. In any case, the growing availability of large datasets on stand biomass and individual plant construction  \citep{Falster2015} suggest the time is ripe for rigorous benchmarking \citep{abramowitz_towards_2012, dekauwe2014} of GVMs against empirical data.

>>>>

Overall, our results establish general patterns about plant construction and thus lay an empirical base against which models can be benchmarked. We found that PFTs differed in LMF, and that LMF was proportional to LMA. It thus seems possible to integrate a key leaf trait with whole-plant modelling of biomass partitioning. A recent study compared allocation routines in a number of leading ecosystem models \citep{dekauwe2014}, and recommended constraining allocation by observed biomass fractions instead of using constant allocation fractions. Based on our findings, a first approximation within GVMs would be to assume leaf area to stem cross-sectional area a parameter that does not vary between PFTs (notwithstanding effects of plant height on this variable), and vary biomass allocation accordingly. LMA, already a parameter in most GVMs, then gives the ratio of leaf biomass to stem area. Some models already incorporate a similar algorithm (see Methods~\ref{sec:gvmsummary}), while other algorithms may be tuned to yield similar patterns between PFTs as we have presented. In any case, the growing availability of large datasets on stand biomass and individual plant construction  \citep{Falster2015} suggest the time is ripe for rigorous benchmarking \citep{abramowitz_towards_2012, dekauwe2014} of GVMs against empirical data.
